Segmentation Using Attributes

### 2010-04-30 trunk/paper.tex
paper with additive prune

%-------------------------------------------------------------------------------
% technical approach
%-------------------------------------------------------------------------------
\section{Technical approach}
\label{sec:technical}
Our approach begins with building a segmentation tree.  We use the segmentation
approach of \cite{arbelaez09} (in fact, we use their code).  This is a hierarchical
segmentation approach which first detects contours using the \emph{gPb} detector
\cite{maire08}.  Once the contours are detected, regions are discovered using a
watershed algorithm after which an Ultrametric Contour Map (UCM) is constructed,
which defines the region hierarchy.

The top node of this hierarchy is the image itself.  Our algorithm first predicts
attributes on the image.  Our attributes are trained using images from the pascal 
2008 dataset using the same annotations used in \cite{farhadi09}.  We trained using
an SVM with linear kernel.

TO TALK ABOUT:
? Discovered that our implementation learns largely on the object and not truly
on the different attributes.

We will assume that we know that an object is present in the image, as is done in \cite{borenstein04}, as opposed to no knowledge, as in \cite{shotton06}.  So we will have to use tagged examples.

The labeling will be binary - imageground and background, as opposed to multi-label results.

We'll use a conditional Markov Random Field (CRF) similar to \cite{gould08}.  The CRF will encode the neighborhood structure/adjacency naturally.

We need to decide whether to use superpixels, as in \cite{gould08} or bag of regions as in \cite{gu09, russell06}, and whether to maintain hierarchy or not.  Another approach to consider is in \cite{hoiem05}.

We have to be careful on attribute definitions, to decide exactly where they lie on the feature spectrum.  I think we should restrict ourselves to labeling cars, and thus the attributes will be things like wheel, door, window, headlight, running board, taillight, etc.

%\begin{figure}
%\center
%\includegraphics[width=0.80\textwidth]{flowchart.eps}
%\caption{Flow chart of testing phase of our algorithm}
%\label{fig:flow}
%\end{figure}
%
%Figure \ref{fig:flow} shows the basic elements of the testing phase of our algorithm.  
%We begin by hierarchically segmenting the image, retaining the hierarchy of the
%segmentation.  Each segment, or region, is then assigned attributes using the algorithm
%found in \cite{farhadi09}.  We will use the code for this algorithm that is 
%provided by the authors.  
%
%Once we have attributes assigned to each region we set up
%the CRF.  The energy function will look something like this:
%
%\begin{equation}
%\psi(\mb{m}) = \sum_i{\left(\phi(I|m_i)+ \sum_j{f^{\mathsf{pair}}(m_i,m_j,I)} \right)}
%\end{equation}
%
%where
%
%\begin{equation}
%\phi(I|m_i) = f^{\mathsf{relloc}}(S_i, c_i, I) + f^{\mathsf{app}}(S_i, c_i, I)
%\end{equation}
%
%$f^{\mathsf{relloc}}(S_i, c_i, I)$ and $f^{\mathsf{app}}(S_i, c_i, I)$ are defined much in 
%the same spirit as they are in \cite{gould08}.  The relative location feature 
%function requires relative location probability maps which we may build by hand
%or learn (see Tasks section below).  The features will be built from the probability
%maps using a voting scheme as in \cite{gould08}.
%
%We are still unsure what the smoothness function 
%$f^{\mathsf{pair}}(m_i,m_j,I)}$ will look like.  We may learn pairwise attribute
%affinity weights or simply determine a weight based on how many car attributes
%the two regions $S_i$ and $S_j$ have.
%
%As we've chosen to do a binary segmentation (figure, ground) we can use a min-cut-based
%algorithm \cite{boykov01} for the energy minimization of the CRF.  We hope to be able
%to use existing code for this (we can't use the code in \cite{gould08} as they don't
%use min-cut), but have not yet looked carefully.
%
%Once the CRF is in place, we run the energy minimization which gives us the
%segmentation with the estimated highest posterior (if we understand it
%correctly).

%-------------------------------------------------------------------------------
% experimental results
%-------------------------------------------------------------------------------
\section{Experimental results}
\label{sec:results}
Here are some results.

>>>>

%-------------------------------------------------------------------------------
% technical approach
%-------------------------------------------------------------------------------
\section{Attribute Detection}
\label{sec:attribute}

Attributes are high-level features which can reflect some semantic properties of objects. Given the attributes, we can do object detection based on the their distribution. We use the same attributes set as \cite{farhadi09}. This set contains three types of attributes, including Shape, Part and Material. All the attributes are listed in figure ?.

We use similar algorithm as \cite{farhadi09} to train the attribute models. Three different type of features are used, such as HOG, texture and color descriptors\cite{farhadi09}. Those features can capture different properties. We use kmeans to quantize all the features. The HOG descriptors are quantized into 1000 clusters, and the texture and color are quantized to 256 and 128 clusters, respectively. All the quantized features are concatenated to form the 1384 deminsional feature.

For training, we use SVM classifier with Gaussian kernel to train model for each attribute. The precision of all the attributes on the test set is illustrated in figure?.

%Our approach begins with training attribute classifiers.  We trained using
%an SVM with linear kernel.  A set of training images is annotated with objects
%contained in the image, bounding boxes for the objects, and attributes exhibited
%by the image.  We use a subset of the features used in \cite{farhadi09}, namely
%HOG, texture and color descriptors.  The HOG descriptors are quantized into 1000
%kmeans centers, and the texture and color are quantized to 256 and 128 kmeans centers,
%respectively.  Our features differ from \cite{farhadi09} in various ways.  We don't
%use edge descriptors, nor do we generate ``sub-''histograms for 6 cells in the
%image.  The presence of cell histograms would complicate matters when extracting
%features from segmented regions.

\section{Hierarchy Segmentation}
\label{sec:segmentation}

We use the segmentation
approach of \cite{arbelaez09}.  This produces a hierarchical
segmentation by first detecting contours using the \emph{gPb} detector
\cite{maire08}.  Once the contours are detected, regions are discovered using a
watershed algorithm after which an Ultrametric Contour Map (UCM) is constructed,
which defines the region hierarchy. Some example of the hierarchy segmentation is shown in figure?

After the construction of hierarchy segmentation tree, we detect all the attributes in each node, including the intermedia node and leaf node. We use the output of SVM as the confidence (not binary), so each node can be represented by the confidence distribution of all attributes.

\section{Object Detection and Segmentation}
\label{sec:detection}

Our objective is to detect object and segment it out. In our hierarchy segmentation tree, the problem changes to how to select those nodes belong to the desired object. We estimate the probability that a node is belong to object based on the attribute confidence distribution. For each object, we manually give the desired attribute distribution by common knowledge. We use the histogram intersection between the desired distribution and the estimated confidence distribution in each node as the similarity measure. The histogram intersection is calculated as follows:

\[ HI(H_{model},H_{pred}) = \frac{|H_{model} \cap H_{pred}|}{|H_{model}|} \]

To select the nodes, we propose two different top-down prune algorithms, additive algorithm and subtractive algorithm. We use the UCM to generate two types of region tree, shown in figure \ref{fig:tree}.
The first tree is an additive tree, where regions are subdivided into successively
smaller regions moving down in the tree.  We call it ``additive'' because the
original image can be reconstructed by adding the regions from each leaf node
together.  The figures show the nodes of the trees as containing images, but in
truth the trees contain binary masks which, when applied to the original image,
produce the shown node contents.

The second tree is ``subtractive.''  The tree has identical structure to the additive
tree and each node contains the inverse of the mask of the corresponding node
in the additive tree.

The final segmentation is obtained by pruning the region tree until a final segmentation
meeting some criterion of absolute or relative attribute saliency is found.
We have two methods for pruning the tree, each showing relative strengths and
weakness.  The first is a semi-greedy additive approach which uses the additive
tree.  The second is a greedy subtractive approach.  We now discuss each of these 
in detail.

\begin{figure*}
\label{fig:tree}
\centering
\subfigure[]{
\label{fig:treea}
\includegraphics[width=0.45\textwidth]{figures/tree_additive.eps}
}
\subfigure[]{
\label{fig:treeb}
\includegraphics[width=0.45\textwidth]{figures/tree_subtractive.eps}
}
\caption{The two segmentation hierarchies used by the additive \subref{fig:treea}
and subtractive \subref{fig:treeb} algorithms.  The additive hierarchy breaks the
regions into smaller and smaller regions, while the subtractive hierarchy increases
the size of the regions as smaller and smaller chunks are removed from the image.}
\end{figure*}

\subsection{Additive algorithm}
\label{ssec:additive}

This prune algorithm is based on three observations. (1) At least one node in the segmentation tree contrains the whole object. (This is always true since the root node contains the whole image). (2) If both parrent node and child node have high confidence as the object, the segment of child node is more accurate than that of the parrent node. (3) Even if the some node has high confidence as the object, its segmentation still need to improve. The details of the additive algorithm is shown in alg ?.\\

The idea of this algorithm is to find out the minimum discriminative segments, either positive or negative, and combine all of them together to form the final segment. We give three threshold for the similarity measure, $thre\_perf$, $thre\_pos$ and $thre\_neg$ which represent perfect segment, positive segment and negative segment, respectively. If the similarity is above $thre\_perf$, which means the segment is a perfect object, we cut all the nodes below and return this segment. If the similarity is below $thre\_neg$, which means the segment is impossible to be related to the object, we cut all the nodes below too and return NULL. If the similarity is between $thre\_perf$ and $thre\_pos$, which means the node definitely contains the object but the segmentation is not perfect yet, we keep on checking its children nodes to refine the segmentation. The final case is that the similarity is between $thre\_pos$ and $thre\_neg$, which means we are not sure whether there is an object in the node, so we just keep on searching its children. \\

The object detection is based on the output of ADDITIVE\_PRUNE. If the output is not NULL, which means at least one node in the tree has similarity above $thre\_pos$, we think the object is detected in this image.\\

(We also plan to do the vertical confidence propogation. But the experiment is too slow, we cannot incorporate it in this draft. We will show it in the final paper.)\\

\begin{algorithm}
\begin{algorithmic}
  \STATE \textbf{input:} 
  \STATE \hspace{3 mm} node $n$ in segmentation tree
  \STATE \hspace{3 mm} desired object model $obj\_model$
  \STATE \hspace{3 mm} threshold for perfect segment $thre\_perf$
  \STATE \hspace{3 mm} threshold for positive segment $thre\_pos$
  \STATE \hspace{3 mm} threshold for negative segment $thre\_neg$
  \STATE \textbf{output:} 
  \STATE \hspace{3 mm} refined segment 
  \STATE
  \STATE $sim=HI(obj\_model,n.conf)$
  \IF{$sim \ge thre\_perf$}
    \STATE return $n.segment$
  \ENDIF
  \IF{$sim \le thre\_neg$}
    \STATE return $NULL$
  \ENDIF
  \STATE $s'=NULL$
  \FOR{$i=1$ to $n.childNum$} 
    \STATE s\{i\}=ADDITIVE\_PRUNE(n.children\{i\},obj\_model,...
    \STATE ... thre\_perf,thre\_pos,thre\_neg)
    \STATE s'=s'+s\{i\}
  \ENDFOR
  \IF{$sim\ge thre\_pos$ and $s'==NULL$}
    \STATE return n.segment
  \ELSE
    \STATE return s'
  \ENDIF
\end{algorithmic}
\caption{ADDITIVE\_PRUNE}
\label{alg:add_prun}
\end{algorithm}

\subsection{Subtractive algorithm}

The subtractive algorithm relies on the idea that if removing a region from
an image doesn't significantly affect the probability of the image containing
the object (which probability is obtained using the predicted attribute 
distribution) then that region can be safely removed in the final segmentation.
The algorithm itself is shown in algorithm \ref{alg:sub_prune}.

\begin{algorithm}
\begin{algorithmic}[1]
  \STATE \textbf{input:} 
  \STATE \hspace{3 mm} node $n$ in segmentation tree
  \STATE \hspace{3 mm} master mask $m_m$
  \STATE \hspace{3 mm} probability $P_{root}$ on root node
  \STATE \hspace{3 mm} factor $f$
  \STATE $m_n$ := mask of $n$
  \STATE $H_{pred}$ := predict attributes on image($m_n \cap m_m$)
  \STATE $\displaystyle P_{pred} = \frac{H_{pattern} \cap H_{pred}}{H_{pattern} \cup H_{pred}}$
  \IF{$P_{pred} < f P_{root}$}
    \STATE SUBTRACTIVE\_PRUNE(n.children)
  \ELSE
    \STATE $m_m = m_m \cap m_n$
  \ENDIF
\end{algorithmic}
\caption{SUBTRACTIVE\_PRUNE}
\label{alg:sub_prun}
\end{algorithm}

The subtractive algorithm begins at the root of the tree and predicts attributes.
This attribute distribution and its associated probability $P_{root}$ is assumed to be
canonical for the image.  That is, any final segmentation should not have
significantly less probability $P_{pred}$ than $P_{root}$.  The algorithm traverses
the subtractive tree depth-first, predicting on each node.  
Any node that doesn't have a significant negative effect on the probability
of the presence of the object is removed.  This test is done on line 9.  If the
predicted probability was significantly affected, then we know that the region
must remain.  But it doesn't mean that sub-regions might be able to be pruned
out.  So the procedure is called recursively, checking for smaller regions
that might be able to be pruned (line 10).

If the probability was not significantly affected, then the mask is applied
(line 12) and no recursive call is made.  This is important, because it means
that we can remove large chunks of the image wholesale and bypass predicting
on the node's posterity.

$f$ is an important parameter that requires tuning.  Increasing $f$ makes the
algorithm more conservative, that is, less likely to apply removals.  Decreasing
$f$ makes it more liberal.  We found $f=0.9$ to be a reasonable value for most images.
Figure \ref{fig:f} shows results of adjusting $f$.

The runtime performance of the algorithm is dominated by the prediction step.
Thus the advantage of pruning large chunks early is highly beneficial to 
the runtime of the algorithm.

\begin{figure}
\centering
\subfigure[]{
\label{fig:fa}
\includegraphics[width=0.22\textwidth]{figures/parking_0.9.eps}
}
\subfigure[]{
\label{fig:fb}
\includegraphics[width=0.22\textwidth]{figures/parking_1.0.eps}
}
\caption{Result of adjusting the factor $f$ in the subtractive algorithm.
\subref{fig:fa} shows running the algorithm with $f=0.9$ and \subref{fig:fb}
shows running the algorithm with $f=1.0$.  Note that \subref{fig:fb} retains
more of the cars, especially the two closest on the right and left, but also
hsa more false-positive pixels, such as the ceiling region of the parking
garage.}
\label{fig:f}
\end{figure}

%-------------------------------------------------------------------------------
% experimental results
%-------------------------------------------------------------------------------
\section{Experimental results}
\label{sec:results}

In this section, we evaluate our attribute based object detection and segmentation algorithm. We will answer four questions. (1) What is the performance of object detection and segmentation? (2) What is the influences of different parameters? (3) What are the advantages of the two pruning algorithms? (4) Is vertical confidence propogation useful?

\subsection{Dataset}
\label{ssec:dataset}

We use Pascal2008

Here are some results.
 trained using images from the pascal 
2008 dataset using the same annotations used in \cite{farhadi09}.
